---
title: "HW 03"
author: "Joey Garcia"
format:
  html:
    embed-resources: true
toc: true
---

```{r setup and load libraries}
if (!require("pacman")) 
  install.packages("pacman")

pacman::p_load(tidyverse, lubridate, glue, scales, dplyr, ggthemes, # most necessary libs.
               forcats,  # reorder function
               ggimage,  # change background with image
               fs,       # filesystem lib.
               janitor  # clean data
               )       

# set theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))

# set width of code output
options(width = 65)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        # 7" width
  fig.asp = 0.618,      # the golden ratio
  fig.retina = 3,       # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300             # higher dpi, sharper image
)

```

## 1 - Du Bois challenge.

```{r}
#| label: Du-Bois-Challenge

income <- read.csv("data/income.csv")

# Add "average_income" to Class Labels
income <- income |>
  mutate(
    Other = replace_na(Other, 0),
    income_label = paste0("$", format(Average_Income, big.mark = ",")),
    ClassLabel = paste0(Class, " | ", income_label),
    ClassLabel = factor(ClassLabel, levels = unique(ClassLabel))
  )

# Define Category order and colors
ordered_categories <- c("Other", "Tax", "Clothes", "Food", "Rent")

category_colors <- c(
  Rent = "#161616",
  Food = "#7d6683",
  Clothes = "#b78a77",
  Tax = "#a9a09d",
  Other = "#bdb09f"
)

# Prepare income data for bar plot
income_clean <- income |>
  select(ClassLabel, Rent, Food, Clothes, Tax, Other) |>
  pivot_longer(cols = Rent:Other, names_to = "Category", values_to = "Percent") |>
  mutate(Category = factor(Category, levels = ordered_categories)) |>
  group_by(ClassLabel) |>
  mutate(pos = cumsum(Percent) - Percent / 2) |>
  ungroup()


# Plot
income_plot <- income_clean |>
  ggplot(aes(x = fct_rev(ClassLabel), y = Percent, fill = Category)) +
  geom_bar(
    stat = "identity", 
    position = "stack", 
    color = "black",
    width = 0.7
  ) +
  geom_text(
    data = income_clean |> filter(Percent > 1),
    aes(label = paste0(round(Percent, 1), "%"), y = pos),
    color = "white", 
    size = 2
  ) +
  scale_fill_manual(values = category_colors) +
  coord_flip(xlim = c(0, 7)) +
  scale_y_continuous(labels = NULL, expand = expansion(mult = c(0.1, 0.1))) +
  
  # Place Social Labels
  annotate(
    "text", x = 1, y = 102, 
    label = "Well-To-Do", 
    size = 2, angle = 90
  ) +
  annotate(
    "text", x = 2.5, y = 102, 
    label = "Comfortable", 
    size = 2,angle = 90
  ) +
  annotate(
    "text", x = 4.5, y = 102, 
    label = "Fair", 
    size = 2, angle = 90
  ) +
  annotate(
    "text", x = 6.5, y = 102, 
    label = "Poor", 
    size = 2, angle = 90
  ) +
  # Bottom text annotation
  annotate(
    "text", x = 0, y = 45, 
    label = "FOR FURTHER STATISTICS RAISE THIS FRAME.", 
    size = 2
  ) +
  labs(
    x = "",
    y = "",
    fill = "Annual Expenditure For"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_text(face = "bold", size = 7),
    panel.grid = element_blank(),
    legend.position = "top",
    legend.title.position = "top",
    legend.title = element_text(hjust = 0.5),
  ) +
  guides(fill = guide_legend(reverse = TRUE))

# Load background image
income_plot <- ggbackground(income_plot, "images/parchment_paper.jpg")
print(income_plot)

```

## 2 - COVID survey - interpret

### Example 1

For the questions, "Based on my understanding, I believe this vaccine is safe" and "I am concerned about my safety and side effects of the vaccine", the wider error bars indicate higher variability in responses. This may suggest that there are widely differing perceptions of vaccine safety and potential side effects. Based on my intuition, this is a viable sentiment, this may reflect underlying differences in trust, prior knowledge, or exposure to misinformation.

### Example 2

Participants that received the COVID and Flu vaccine expressed a similar sentiment of "feeling safer at work" and confidence of the COVID vaccine vetting process. This shared perspective may reflect institutional trust among dual-vaccinated individuals. In contrast, non-vaccinated participants express skepticism with less confidence in the vaccine development process and reduced trust in the institutions promoting vaccination. I agree with these results, this is highlighting a divergence in vaccines perception of credibility and safety.

### Example 3

Nursing and Medical Participants returned surprising results for "feeling safer at work" and confidence of the COVID vaccine vetting process. Nurses showed low variation in their responses with narrower error bars, indicating a high level of trust. In contrast, medical participants exhibited a wider range of responses with larger error bars suggesting greater variability in confidence and perceived workplace safety. This divergence may point to differing experiences, levels of autonomy, or access to information between the two professional groups.

## 3 - COVID survey - reconstruct

```{r}
#| label: COVID-clean

# Load COVID Survey
survey <- read.csv("data/covid-survey.csv")

# Clean Names and Filter NAs
survey_clean <- survey |> 
  row_to_names(row_number = 1) |>
  clean_names() |>
  mutate(across(everything(), ~ na_if(., ""))) |> # empty strings to NA
  filter(if_any(-response_id, ~ !is.na(.)))       # Remove rows where all except 'response_id' are NA

survey_clean |>
  dim()
```

``` {r}
#| label: COVID-recode

# recode variables
survey_recoded <- survey_clean |>
  mutate(
    exp_already_vax = recode(
      exp_already_vax, "0" = "No", "1" = "Yes"),
    exp_flu_vax = recode(
      exp_flu_vax, "0" = "No", "1" = "Yes"),
    exp_profession = recode(
      exp_profession, "0" = "Medical", "1" = "Nursing"),
    exp_gender = recode(
      exp_gender, 
      "0" = "Male",
      "1" = "Female", 
      "3" = " Non-binary third gender", 
      "4" = "Prefer not to say"
      ),
    exp_race = recode(
      exp_race, 
      "1" = "American Indian / Alaskan Native", 
      "2" = "Asian", 
      "3" = "Black / African American", 
      "4" = "Native Hawaiian / Other Pacific Islander", 
      "5" = "White"
      ),
    exp_ethnicity = recode(
      exp_ethnicity, 
      "1" = "Hispanic / Latino", 
      "2" = "Non-Hispanic/Non-Latino"
      ),
    exp_age_bin = case_when(
      exp_age_bin == "0" ~ "<20",
      exp_age_bin == "20" ~ "20–25",
      exp_age_bin == "25" ~ "26–30",
      exp_age_bin == "30" ~ ">30",
      TRUE ~ NA_character_  # For NA values
      )
  )

survey_recoded |>
  dim()
```

``` {r}
#| label: COVID-summary_stats-data-wrangling

covid_survey_longer <- survey_recoded |>
  pivot_longer(
    cols = starts_with("exp_"),
    names_to = "explanatory",
    values_to = "explanatory_value"
  ) |>
  filter(!is.na(explanatory_value)) |>
  pivot_longer(
    cols = starts_with("resp_"),
    names_to = "response",
    values_to = "response_value"
  )

# create quantile values for high and low probabilities
covid_survey_summary_stats_by_group <- covid_survey_longer |>
  group_by(explanatory, explanatory_value, response) |>
  summarize(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low  = quantile(as.numeric(response_value), probs = 0.1, na.rm = TRUE),
    high = quantile(as.numeric(response_value), probs = 0.9, na.rm = TRUE),
    .groups = "keep"
  ) 

covid_survey_summary_stats_by_group

# introduce "All" as an explanatory row
covid_survey_summary_stats_all <- covid_survey_longer |>
  group_by(response) |>
  summarize(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low  = quantile(as.numeric(response_value), probs = 0.1, na.rm = TRUE),
    high = quantile(as.numeric(response_value), probs = 0.9, na.rm = TRUE),
    explanatory = "All",
    explanatory_value = "",
    .groups = "drop"
  )

covid_survey_summary_stats_all

# Bind summary stats by rows
covid_survey_summary_stats <- bind_rows(
  covid_survey_summary_stats_by_group,
  covid_survey_summary_stats_all
)

covid_survey_summary_stats

```

``` {r}
#| label: COVID-survey-recreation
#| fig-height: 8

response_labels <- c( 
  "resp_safety" = 
    "Based on my understanding, I believe the vaccine is safe",
  "resp_feel_safe_at_work" = 
    "Getting the vaccine will make me feel safer at work",
  "resp_concern_safety" = 
    "I am concerned about the safety and side effects of the vaccine",
  "resp_confidence_science" = 
    "I am confident in the scientific vetting process for the new COVID vaccines",
  "resp_trust_info" = 
    "I trust the information that I have received about the vaccines",
  "resp_will_recommend" = 
    "I will recommend the vaccine to family, friends, and community members"
)

explanatory_labels <- c(
  "All" = "All",
  "exp_age_bin" = "Age",
  "exp_gender" = "Gender",
  "exp_race" = "Race",
  "exp_ethnicity" = "Ethnicity",
  "exp_profession" = "Profession",
  "exp_already_vax" = "Had COVID vaccine",
  "exp_flu_vax" = "Had flu vaccine this year"
)

covid_plot <- covid_survey_summary_stats |>
  # adjust feature and response names
  mutate(
    response = recode(response, !!!response_labels),
    explanatory = recode(explanatory, !!!explanatory_labels),
    explanatory = factor(explanatory, levels = explanatory_labels) # reorder
  ) |>

  ggplot(aes(y = fct_rev(explanatory_value))) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = 0.4) +
  geom_point(aes(x = mean), color = "black", size = 2) +
  facet_grid(
    explanatory ~ response, 
    labeller = labeller(
      response = label_wrap_gen(15),
      explanatory = label_wrap_gen(15)), 
    scales = "free_y"
  ) +
  labs(
    title = NULL,
    x = "Mean Likert Scored\n(Error bars range from 10th to 90th percentile)",
    y = NULL
  ) +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(
    strip.text.y = element_text(angle = 0, hjust = 0.5, size = 12), # Rotate "explanatory" labels
    strip.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "grey90")
  )

covid_plot

```

## 4 - COVID survey - re-reconstruct

```{r}
#| label: COVID-re-reconstruct
#| fig-height: 8


# create quantile values for high and low probabilities
covid_survey_summary_stats_by_group_alt <- covid_survey_longer |>
  group_by(explanatory, explanatory_value, response) |>
  summarize(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low  = quantile(as.numeric(response_value), probs = 0.25, na.rm = TRUE),
    high = quantile(as.numeric(response_value), probs = 0.75, na.rm = TRUE),
    .groups = "keep"
  ) 

# introduce "All" as an explanatory row
covid_survey_summary_stats_all_alt <- covid_survey_longer |>
  group_by(response) |>
  summarize(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low  = quantile(as.numeric(response_value), probs = 0.25, na.rm = TRUE),
    high = quantile(as.numeric(response_value), probs = 0.75, na.rm = TRUE),
    explanatory = "All",
    explanatory_value = "",
    .groups = "drop"
  )

# Bind summary stats by rows
covid_survey_summary_stats_alt <- bind_rows(
  covid_survey_summary_stats_by_group_alt,
  covid_survey_summary_stats_all_alt
)                                       


covid_plot_alt <- covid_survey_summary_stats_alt |>
  # adjust feature and response names
  mutate(
    response = recode(response, !!!response_labels),
    explanatory = recode(explanatory, !!!explanatory_labels),
    explanatory = factor(explanatory, levels = explanatory_labels) # reorder
  ) |>

  ggplot(aes(y = fct_rev(explanatory_value))) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = 0.4) +
  geom_point(aes(x = mean), color = "black", size = 2) +
  facet_grid(
    explanatory ~ response, 
    labeller = labeller(
      response = label_wrap_gen(15),
      explanatory = label_wrap_gen(15)), 
    scales = "free_y"
  ) +
  labs(
    title = NULL,
    x = "Mean Likert Scored\n(Error bars range from 25th to 75th percentile)",
    y = NULL
  ) +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(
    strip.text.y = element_text(angle = 0, hjust = 0.5, size = 12), # Rotate "explanatory" labels
    strip.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "grey90")
  )

covid_plot_alt
```

### Response

When comparing the two graphs, the immediate difference is the smaller intervals because of a reduced percentile range. These percentiles are specified as probabilities between 0 and 1 in the quantile() function. 

My view has changed based on the new plot, read the following:

For example 1, the error bars are more narrow for "Based on my understanding, I believe this vaccine is safe". My conclusion changes, the majority of the error bars now span across the 1 and 2 response values. In contrast, the error bars for the original plot spans across all response values 1 through 5. The new plot portrays to the audience that participants are more certain the "vaccine" is safe. 

No change, for example 2. 

For example 3, similar to example 1, the error bars are more narrow and show less variability in the data. The answers among Nurses and Medical doesn't indicate any sign of divergence in opinions. Based on the new plot, Nurses and Medical now show error bars that trend towards positive sentiment towards "trusting the vetting process" and "safer at work".

## 5 - COVID survey - another view

```{r}
#| label: COVID-alternate-plot



```
